---
title: \normalsize Title
csl: Bibliography/anesthesia-and-analgesia.csl
fontfamily: pslatex
output:
  pdf_document: default
---

\setlength{\parindent}{0ex}

##### Authors

Michael H. Andreae, M.D., Department of Anesthesiology, Penn State Health Milton S. Hershey Medical Center, Penn State College of Medicine, Hershey. PA

Nathan Pace, PhD., Department of Anesthesiology, Univeristy of Utha, Salt Lake City, UT

##### Corresponding author:
Michael Andreae, MD, Address: Department of Anesthesiology, Department of Anesthesiology, Penn State Health Milton S. Hershey Medical Center, 500 University Drive, Hershey. PA 17033, Tel: +1 (717) 531 6140, Email: mandreae@pennstatehealth.psu.edu

##### Acknowledgement

We would like to acknowledge Doleman et al for providing the data to generate the figures.

##### Conflicts of Interest: None

##### Keywords:
analgesic adjuvants, postoperative pain, evidence synthesis, meta-regression, Bayesian statistics


\newpage
\setlength{\parindent}{0ex}


```{r, global_options, echo=FALSE}
suppressWarnings(library(knitr))
# set options
opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
               echo=FALSE, warning=FALSE, message=FALSE)
eval_everything=FALSE
```


\newpage

\setlength\parindent{0pt}

# Editorial



# Figures

```{r GlobalLibraries}
library(ggplot2)
```

```{r ImportData, eval=TRUE}
library(readxl) # read excel file

Acetaminophen <- read_excel("data/Acetaminophen.xlsx")
Acetaminophen <- Acetaminophen[-1,] #drop the empty first line
Acetaminophen$Adjuvant <- rep("Acetaminophen",dim(Acetaminophen)[1])
save(Acetaminophen, file = "data/Acetaminophen.Rdata") #drop the empty first line

NSAIDS2 <- read_excel("data/NSAIDS2.xlsx")
NSAIDS2 <- NSAIDS2[-1,] #drop the empty first line
NSAIDS2$Adjuvant <- rep("NSAIDS2",dim(NSAIDS2)[1])
save(NSAIDS2, file = "data/NSAIDS2.Rdata")

Alpha_2 <- read_excel("data/Alpha-2.xlsx")
Alpha_2 <- Alpha_2[-1,] #drop the empty first line
Alpha_2$Adjuvant <- rep("Alpha_2",dim(Alpha_2)[1]) 
save(Alpha_2, file = "data/Alpha_2.Rdata")

## Combine all three sets into one
AllAdjuvants <- rbind(Acetaminophen, NSAIDS2, Alpha_2)
# factorize surgery type and adjuvant 
AllAdjuvants$Surgery <- as.factor(AllAdjuvants$`Type of surgery`)
AllAdjuvants$Agent <- as.factor(AllAdjuvants$Adjuvant)

save(AllAdjuvants, file= "data/AllAdjuvants.Rdata")
```

```{r LoadData, eval=TRUE}
rm(list=ls())
load(file = "data/Alladjuvants.Rdata")
str(AllAdjuvants)
```

## Figure 2

```{r DraftFigure}
# assign only the Acetaminophen data to ggplot object ggData
ggData <- ggplot(AllAdjuvants[AllAdjuvants$Agent=="Acetaminophen",], aes(x=X,y=Y, colour= Surgery)) 
# make a Scatter plot and assign it to Fig
Fig_a <- ggData +geom_point() 

## add the regression line
Fig_b <- Fig_a + geom_abline(intercept = 0, slope = (-18.67/50), colour="red", size = 1)

# choose a simple background and render
Fig_b + theme_bw() # choose a simple background and render
```

### Figure 2 Caption
In Figure 2, we reproduce the acetaminophen subplot of Figure 2 of Doleman et al. Each study is represented by a dot. We colored the studies by surgical procedure, according to the adjacent colour legend. If surgery were the best way to explain why studies yield different results, then studies investigating the same surgical procedure should have similar results. Hence the same color dots should be clumped together,(and the green Cholecystectomy studies are somewhat close).  But most studies of the same color (investigating the same procedure) are far apart, for example orthopedic (blue) or hysterectomy (violett) study dots are spread out over the entire range of the plot. By contrast, all studies across all surgery types line up neatly around the red regression line, when we plot the mean reduction in morphine consumption (y-axis) against the baseline risk (baseline morphine consumption in the control group). The conclusion is that a meta-regression controlling for baseline morphine consumption in the control group is superior to stratification by surgery in explaining between study variance in results.
