---
title: \normalsize Title
csl: Bibliography/anesthesia-and-analgesia.csl
fontfamily: pslatex
output:
  pdf_document: default
---

\setlength{\parindent}{0ex}

##### Authors

Michael H. Andreae, M.D., Department of Anesthesiology, Penn State Health Milton S. Hershey Medical Center, Penn State College of Medicine, Hershey. PA

Nathan Pace, PhD., Department of Anesthesiology, Univeristy of Utha, Salt Lake City, UT

##### Corresponding author:
Michael Andreae, MD, Address: Department of Anesthesiology, Department of Anesthesiology, Penn State Health Milton S. Hershey Medical Center, 500 University Drive, Hershey. PA 17033, Tel: +1 (717) 531 6140, Email: mandreae@pennstatehealth.psu.edu

##### Acknowledgement

We would like to acknowledge Doleman et al for providing the data to generate the figures.

##### Conflicts of Interest: None

##### Keywords:
analgesic adjuvants, postoperative pain, evidence synthesis, meta-regression, Bayesian statistics


\newpage
\setlength{\parindent}{0ex}


```{r, global_options, echo=FALSE}
suppressWarnings(library(knitr))
# set options
opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
               echo=FALSE, warning=FALSE, message=FALSE)
eval_everything=FALSE
```


\newpage

\setlength\parindent{0pt}

# Editorial



# Figures

```{r GlobalLibraries}
library(ggplot2)
```

```{r ImportData, eval=TRUE}
library(readxl) # read excel file

Acetaminophen <- read_excel("C:/Users/Michael/Downloads/Acetaminophen.xlsx")
Acetaminophen <- Acetaminophen[-1,] #drop the empty first line
save(Acetaminophen, file = "data/Acetaminophen.Rdata") #drop the empty first line

NSAIDS2 <- read_excel("C:/Users/Michael/Downloads/NSAIDS2.xlsx")
NSAIDS2 <- NSAIDS2[-1,] #drop the empty first line
save(NSAIDS2, file = "data/NSAIDS2.Rdata")

Alpha_2 <- read_excel("C:/Users/Michael/Downloads/Alpha-2.xlsx")
Alpha_2 <- Alpha_2[-1,] #drop the empty first line
save(Alpha_2, file = "data/Alpha_2.Rdata")
```

```{r LoadData, echo=TRUE}
rm(list=ls())
load(file = "data/Alpha_2.Rdata")
load(file = "data/NSAIDS2.Rdata")
load(file = "data/Acetaminophen.Rdata")

str(Acetaminophen)
```

## Figure 1

```{r DraftFigure}
Fig <- ggplot(Acetaminophen, aes(x=X,y=Y))
Fig+geom_point()
```

### Figure 1 Caption

## Figure 2

### Figure 2 Caption

